---
title: "Putting Terrorism In Context"
date: "2023-04-19"
output:
  html_document:
    number_sections: false
    toc: true
    fig_width: 8
    fig_height: 6
    theme: cosmo
    highlight: tango
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install required libraries

```{r echo=TRUE}
library(ggplot2)
library(treemap)
library(corrplot)
library(tm)
library(tidytext)
library(tidyr)
library(wordcloud)
library(knitr)
library(kableExtra)
library(formattable)
library(dplyr)
library(topicmodels)
library(readr)
library(lubridate)
library(dplyr)
library(magrittr)
library(leaflet)

options(knitr.table.format = "html")
```

```{r echo=TRUE}
terror <- read.csv('../globalterrorismdb_0718dist.csv', header = TRUE, stringsAsFactors = FALSE, na.strings = c("", "Unknown"))
```


```{r echo=TRUE}

KeepCols <- c('eventid','iyear', 'imonth', 'iday','extended','country_txt','region_txt','provstate','city','latitude','longitude','summary','alternative_txt','success','attacktype1_txt','targettype1_txt','targsubtype1_txt','natlty1_txt','gname','motive','nperps','weaptype1_txt','nkill')

```


```{r echo=TRUE}

terror_df <- terror[colnames(terror) %in% KeepCols]
terror_df$iday[terror_df$iday == 0] <- 1 #there are some days with value zero. making it into 1
terror_df$imonth[terror_df$imonth == 0] <- 1 
terror_df$iyear <- as.factor(terror_df$iyear)
terror_df$imonth <- as.factor(terror_df$imonth)
terror_df$iday <- as.factor(terror_df$iday)
terror_df$date <- as.Date(paste(terror_df$iyear, terror_df$imonth, terror_df$iday, sep = "-"), format = "%Y-%m-%d")

```

```{r}
colnames(terror_df)

```


#Let us check how the terrirism events have increased over the years

```{r}
# group the data by year and count the number of events
events_per_year <- terror_df %>%
  group_by(year = year(date)) %>%
  summarize(num_events = n())


# create a bar chart showing the number of events per year
ggplot(events_per_year, aes(x = year, y = num_events, fill = num_events)) +
  geom_bar(stat = "identity") +
  scale_fill_gradient(low = "lightpink", high = "red") +
  labs(x = "Year", y = "Number of events", title = "Number of terrorist events over time")

```
Terrorist attacks have increased tremendously from 2012, peaking to ~16k attacks happening that year 

We can see that terrorism started declining after around 1990s towards 2006, but started gaining momentum once again. Why?
Let us try to see what changed in terms of the areas that were targeted the most, the kind of attacks that happened during that period, which insurgent groups emerged and what kind of weapons did terrorists start using.


```{r}
terror_df_pre2006 <- terror_df %>% filter(year(date) < 2006)
terror_df_post2006 <- terror_df %>% filter(year(date) > 2006)
  
```


```{r}

attacks_post2006_by_country <- terror_df_post2006 %>%
  group_by(country_txt) %>%
  summarise(total_attacks = n()) %>%
  arrange(desc(total_attacks)) %>%
  head(15)

# Create a bar plot using ggplot2 and sort by value
ggplot(attacks_post2006_by_country, aes(x = reorder(country_txt, total_attacks), y = total_attacks)) +
  geom_bar(stat = "identity", fill = "red") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  xlab("Country") +
  ylab("Number of attacks") +
  ggtitle("Total Terror Attacks Post-2006 by Country") +
  coord_flip()  # Flip the axis to sort by value


```

```{r}

library(maps)

# Count the number of attacks per country and create a new data frame
attacks_post2006_by_country <- terror_df_post2006 %>%
  group_by(country_txt) %>%
  summarise(total_attacks = n()) %>%
  arrange(desc(total_attacks)) %>%
  slice(1:20) %>%
  drop_na()

# Load the world map data
world_map <- map_data("world")

# Merge the number of attacks data with the world map data
map_data <- merge(world_map, attacks_post2006_by_country, by.x = "region", by.y = "country_txt")

# Create the map plot using ggplot2
ggplot(data = map_data, aes(x = long, y = lat, group = group, fill = total_attacks)) +
  geom_polygon(color = "black") +
  scale_fill_gradient(low = "white", high = "darkred", na.value = "grey90") +
  labs(title = "Total Attacks Post-2006 by Top 20 Countries", fill = "Total Attacks")



```


Which are the countries where most people are killed every year? And what are some of the most affected states in those countries
```{r}


# Calculate the total number of people killed per year for each country and state
kills_by_year_by_country <- terror_df_post2006 %>%
  group_by(iyear, country_txt) %>%
  summarise(total_kills = sum(nkill, na.rm = TRUE))



# Create line plot
ggplot(kills_by_year_by_country, aes(x = iyear, y = total_kills, group = country_txt)) +
  geom_line() +
  labs(x = "Year", y = "Total Kills", color = "Country")



```





